The mass distribution of companions to low-mass white dwarfs

### 2014-11-21 documents/letter/response.txt

Dear Dr. de Grijs,

We thank the two referees for their detailed comments.

We have updated the manuscript based on their suggestions. Below we detail how we addressed the referees' comments.

Please let us know if further improvements are required.

@@

Sincerely,
Jeff J Andrews

>========================================================
>Report from Referee 1:
>
>- Title. Since the sample of study is the ELM sample (defined as
>Mwd<~0.3Msun), instead of the low-mass sample (defined as
>Mwd<~0.45Msun), change the title to "The mass distribution of
>companions to extremely low-mass white dwarfs".

Although the ELM survey is a search for WDs with masses <~0.3 Msun, a significant fraction (~20%) of the sample is composed of WDs with masses from 0.3-0.45 Msun. As a result we think it is more appropriate to keep "low-mass white dwarfs" in the title, particularly since for our purposes this distinction between low-mass and extremely low-mass is not critical.

>- Abstract. "We then apply this model to a sample of low mass white
>dwarfs (LMWDs, M<~0.45 Msun)." Reading this gives the impression that
>the sample is an observed (real) one, which is not the case. Specify
>that the low-mass WD samples used are three, and that these are drawn
>from mock data.

The abstract now reads "We apply this model to three mock samples of low-mass white dwarfs (LMWDs, $M\lesssim0.45~\Msun$) and a sample of post-common-envelope binaries." The test using post-common-envelope binaries (PCEBs) was added in response to a later request from the referee.

>- Introduction. The authors cite the seminal paper by Marsh et
>al. (1995) that demonstrated that a large fraction of low-mass white
>dwarfs are in binaries. However, several more and recent observational
>efforts have been performed that prove to high extent that indeed
>low-mass white dwarfs are formed due to binary interactions. Add 
>citations to e.g. Maxted et al. 2000MNRAS,319,305M, Rebassa-Mansergas 
>et al. 2011MNRAS,413,1121R.

Because of space restrictions, we had to restructure the text somewhat to include the test with the PCEBs. So the introduction reads slightly differently; the relevant text is now "Indeed, with few exceptions, follow-up observations consistently find companions to LMWDs \citep{marsh95,maxted00,nelemans05,rebassa11}." In addition to the citations mentioned by the referee, we include a reference to the SPY survey, which identified two LMWDs in close binaries with unknown companions.

>- Introduction. The authors should state more clearly why identifying
>NS or WD companions to ELM white dwarfs is important. What important
>problems will be solved or can be addressed by knowing that NS are or
>are not likely to be the companions of ELM white dwarfs?

In the second paragraph of the introduction, we have added a sentence explaining the importance of finding even one NS companion to a spectroscopically characterized LMWD: "Finding even one NS companion to a spectroscopically characterized LMWD would be very valuable, since this system could constrain the NS mass."

>- Section 2. Why using a logarithmic prior for sigma(WD)?

We choose a logarithmic prior for sigma(WD) since it is scale-invariant. We have added this to the text, in the second paragraph of Section 2.1.

@@

>Why fixing sigma(NS) to 0.05Msun when sigma(WD) has a much wider range? Is it because the NS mass distribution peaks at ~1.35Msun with almost no tail towards lower and higher masses?

We wanted to build a simplistic model, with few parameters, and therefore set the NS distribution without allowing it to vary. From Kiziltan et al. (2013) and Smedley et al. (2014) it appears that the mass distribution of NSs with WD companions may have a slightly higher-mass tail. We therefore set the NS mass distribution slightly higher than the canonical value of 1.35 Msun. It is not observed to vary much from this, so we made the distribution narrow by setting sigma(NS) to 0.05 Msun.

@@

>According to the text, sigma(WD) ranges from 0.02-2Msun. However, in Table 1 I see sigma(WD) proportional to sigma^(-1), where 0.02 
>- Table 1. The prior of mu(WD) is written as U(0.2,1), but should not it be U(0.2,1.4)?

The prior on sigma(WD) should be U(0.2,2.0) and on mu(WD), 0.02 < mu < 1.0. Both the text and table have been fixed to reflect this. For the flexibility of our model, we allow sigma(WD) to vary greatly. We still truncate the mass distribution at [0.2,1.44], but we do not allow the center of the Gaussian, mu(WD), to go above 1.0 Msun.

>- Figure 1. Why not adding the number and label in the y axis? Top
>left panel. It is interesting to see that the distribution is
>bi-modal, and that there is a large number of massive (~1Msun) white
>dwarfs. How realistic are the assumptions for this particular M2
>distribution?

We have added labels and numbers to the y axes of our M_2 distributions. The distribution in Figure 1 was drawn from a Gaussian with mu(WD)=0.7 Msun and sigma(WD)=0.2. We tested this by drawing other random samples; the bimodal distribution is due to the randomness of drawing 100 WD samples. In other randomly drawn samples, this distribution appears as a single Gaussian. It may not be realistic to have so many WDs with masses above 1 Msun, but it is encouraging that our model can still fairly accurately determine that these systems don’t have NS companions.

>- I consider testing the statistical model using three samples of mock
>data is very useful. However, I believe the model should also be
>tested with real data. The sample of close white dwarf-M dwarf
>binaries with orbital periods and component masses measured is of the
>same order as the ELM survey and can be used for this purpose (see
>e.g. Nebot Gomez-Moran et al. 2011A&A,536A,43N; there are also several
>known eclipsing systems: see the work by S.G. Parsons et al.). In this
>case, K are known for the secondary stars so the distribution to test
>with the statistical model would be the white dwarf mass
>distribution. (Note that for this sample the spectral type of the
>secondary stars are provided, which can be converted into mass
>following a spectral type-mass relation: e.g. Baraffe & Chabrier
>1996ApJ,461L,51B).

Based on this suggestion, we have added a fourth test where we derive the resulting WD mass distribution of companions to the SDSS PCEBs of Nebot Gomez-Moran et al. (2011). Instead of converting the provided spectral types into masses, we include the companion masses provided in this work. Parsons et al. (2012) list two of the systems in this PCEB sample as eclipsing. For one of them, J1348+1834, the value listed differs from the WD mass given by Drake et al. (2010). Because we cannot determine which is the correct mass, we do not include constraints in our analysis based on the fact that these two systems are eclipsing. Since there are only two systems and they do not have extreme masses, we expect that this will not affect the resulting posterior distributions. The description of this test is included in Section 3.4. The histogram of the WD masses in these systems is shown in the bottom left panel of Figure 3. The MAP value shows that our model does a good job of determining the mass distribution of these systems.

>- Section 4.1. "Fitting the WDs' spectra to templates provides
>precisely determined M1 values". I am not entirely convinced by this
>sentence. This is definitely the case when one of the stars is
>considerably more luminous than the other, but it might not be the
>case otherwise. Lets consider a white dwarf which contributes 30% to
>the total flux in the Hbeta to Hepsilon Balmer line regime (where the
>line profile fitting takes place), the other white dwarf contributing
>70%. The total flux will be the sum of the two flux contributions and
>a fit to the resulting spectrum may not give the mass of the more
>luminous white dwarf. I agree that assuming M1 to be the one obtained
>from the template fit is the best way to proceed, however a note of
>caution should be added in the text. A plausible way to check this
>would be to compare the accurate masses of the eclipsing systems with
>those provided from the spectral fitting, if available.

It is possible that companion WDs could contribute to the luminosity of the LMWD primaries. However, Gianninas et al. (2014) point out that because these are all single-lined spectroscopic binaries, they expect the companions' “contribution to the observed spectra to be on the order of a few percent at most.” These authors' extremely high S/N coadded spectra would probably detect spectral lines from any luminous companion. In investigating this possibility, we did find a potentially larger source of uncertainty: for two of the three eclipsing systems, the spectroscopic mass measurements are in agreement with the expectation from the eclipse measurement. However, for the third system, NLTT 11748, there is a significant discrepancy. Gianninas et al. (2014) attribute this difference to the “high log g problem,” one of the possible issues with using 1-D WD atmospheric models. They suggest that only the spectroscopic masses of the seven WDs with temperatures below ~9500 K will be (somewhat) affected. We add a note of caution in the text, in paragraph 2 of Section 3.4, to indicate this effect may lead to some uncertainties in these mass measurements.

>- Section 4.2. "Our distribution is significantly wider (sigma ~ 0.26
>Msun compared with sigma~0.1 Msun), possibly due to mass transfer
>affecting the mass distribution of the unseen primary WDs as well." Do
>the authors mean on-going mass transfer? If that is the case, should
>we not see an accretion disk in the spectrum? It seems unlikely that a
>large fraction of these systems are currently undergoing mass transfer
>episodes. Or do the authors mean previous mass transfer episodes? When
>specifically?

We expect that all these systems are detached due to the lack of emission lines in the spectra, as would be expected from an accretion disk. Paragraph 2 of Section 4.2 has been updated to specify that the companion masses may be more massive due to previous mass-transfer phases.

>- Section 4.2. Since the ELM sample contains primaries of M<~0.3Msun
>and mu(WD)=0.71Msun, this seems to imply that the future mergers will
>not produce type Ia supernovae. I consider the authors should add a
>couple of sentences discussing this.

While most of these systems almost certainly have a combined mass less than 1.4 Msun, many of the combined masses of these systems are still ~1Msun. They may therefore evolve into underluminous supernova. We can only constrain the population; a small subset of LMWDs may have high enough masses to be SNe Ia progenitors. We have included some discussion of this in the third paragraph of Section 4.2.

>- Inspection of figure 3 and the right panel of figure 4 reveals that
>P(NS) increases with mf. Add an explanation why this is the case. Do
>these figures imply that an observer looking for a NS in orbit with a
>ELM white dwarf can focus specifically on targets with the highest mf?
>If so, add a sentence stating this in the text.

Systems with higher mf values are more likely to have NS companions through Equation 1. We added a note about this in the last paragraph of Section 4.2.

>****
>Report from Referee 2:

>A. Bayes' rule is only needed if significant priors are introduced. Are equations (7) and (19)-(20) relevant prior distributions? If so, please label them accordingly. Indeed, the authors state in 4.2/par2 that the "results are independent of astrophysical expectations as they do not include informative priors on ..."; but the entire point of the Bayesian approach is to introduce meaningful priors. Since the authors seek the MAP solution, I suspect the calculation is equivalent to a constrained maximum likelihood calculation (e.g. http://link.springer.com/article/10.1023/A:1008669208700 http://faculty.washington.edu/rons/cmlinf/cmlinf.html). There is no difficulty in using an MCMC approach, and the results are unchanged, but the authors should clarify that the calculation can be viewed from either perspective. Fortunately the authors do not emphasize `Bayesian' in the title, abstract or conclusions.

@@

The prior in equation 7 comes from the fact that the inclination angle is random. One of the advantages to this sample is that there is (nearly) no observational bias on the inclination angle, therefore this prior is appropriate. The priors on our NS model, listed in Equations 19-20, are reasonable based on NS mass measurements derived elsewhere, in Kiziltan et al. (2013) and Smedley et al. (2014).

While we could have approached this problem using a maximum likelihood technique, we would have had to use a regularization scheme to account for the prior on /sigma. Furthermore, we chose a Bayesian approach so that we can make our posterior distributions publicly available. Using these, the NS probabilities can be calculated for newly discovered LMWDs.

>B. Concerning the statement in 4.2 "Our distribution is significantly wider ....[physical explanation]. Couldn't this be a statistical effect: Wouldn't the posterior distribution appear wider just because the sample is small and the spread is thus poorly constrained? This can be checked with simulations of different sample sizes.

In Section 4.2, the statement about our distribution is referring to the companion distribution, not the posterior distribution. Our statistical model should provide an unbiased estimate of \sigma, since it is a model parameter. Therefore, it should not be affected by the sample size.

### 2014-11-21 documents/letter/ms.tex
Updated .tex and response

\title{The mass distribution of companions to low-mass white dwarfs}
\author{Jeff J. Andrews\altaffilmark{\colum}, Adrian M. Price-Whelan\altaffilmark{\colum}, Marcel A. Ag\"ueros\altaffilmark{\colum}}

@@

\begin{abstract}
Measuring the masses of companions to single-line spectroscopic binary stars is (in general) not possible because of the unknown orbital plane inclination.  Even when the mass of the visible star can be measured, only a lower limit can be placed on the mass of the unseen companion. However, since these inclination angles should be isotropically distributed, for a large enough, unbiased sample, the  companion mass distribution can be deconvolved from the distribution of observables. In this work, we construct a hierarchical probabilistic model to infer properties of unseen companion stars given observations of the orbital period and projected radial velocity of the primary star. 
We then apply this model to three mock samples of low mass white dwarfs (LMWDs, $M\lesssim0.45~\Msun$) and a sample of post-common envelope binaries (PCEB). We use a mixture of two Gaussians to model the WD and neutron star (NS) companion mass distributions. Our model successfully recovers the initial parameters of these test data sets. 
%We focus on low mass white dwarfs (LMWDs, $M\lesssim0.45~\Msun$), modeling the companion mass distribution with a mixture of two Gaussians to represent a WD and neutron star (NS) companion population. Our model successfully recovers the initial parameters of three mock samples of low mass white dwarfs (LMWDs, $M\lesssim0.45~\Msun$) and a test sample of post-common envelope binaries (PCEB). 
We then apply our model to 55 WDs in the extremely low mass (ELM) WD Survey. Our maximum a posteriori model for the WD companion population has a mean mass, $\mu_{\rm WD}$, of $0.74~\Msun$ with a standard deviation, $\sigma_{\rm WD}$, of $0.24~\Msun$. Our model constrains the NS companion fraction, $f_{\rm NS}$, $<16$\% at 68\% confidence. We make samples from the posterior distribution publicly available so that future observational efforts may compute the NS probability for newly discovered LMWDs.
\end{abstract}

@@

\section{Introduction}

Except in cases of extreme metallicity \citep{kilic07}, the Galaxy is not old enough to produce LMWDs through single-star evolution. Instead, they form through interactions with another star and are expected to be in binary systems \citep{han98,nelemans00,nelemans01,vdSluys06,woods12}. 
%In their seminal work, \citet{marsh95} discovered companions to five LMWDs, confirming these theoretical expectations. T
Follow-up observations consistently find companions around LMWDs with few exceptions \citep{marsh95,maxted00,nelemans05,rebassa11}. Recently, the ELM WD Survey has searched for extremely LMWDs ($M\lesssim0.3~ \Msun$) in the Sloan Digital Sky Survey \citep[SDSS;][]{york00} and elsewhere \citep{ELMI,ELMII, ELMIII, ELMIV, ELMV}. Fifty-five of their 61 identified LMWDs have a measured radial velocity (RV) and orbital period ($\period$); we refer to these 55 systems as the ELM sample.
% \citep{ELMI} identifies LMWDs in the Sloan Digital Sky Survey \citep[SDSS;][]{york00} and elsewhere for spectroscopic follow-up, focusing on extremely low mass WDs ($M\lesssim0.3~ \Msun$). After these authors' initial investigation found that 11 of 12 LMWDs had close binary companions, further spectroscopic follow-up has now brought the sample of ELM WDs to 61 systems, 55 of which have measured radial velocity amplitudes and orbital periods \citep{ELMII, ELMIII, ELMIV, ELMV}. We refer to these 55 systems as the ELM sample.

RV and $\period$ measurements indicate that the LMWDs companions are most likely WDs. However, since the inclination angle ($i$) is unknown, LMWDs could have NS companions \citep{vLeeuwen07}. Indeed, LMWDs are observed as companions to millisecond pulsars \citep{vKerkwijk96,callanan98,bassa06,antoniadis12}. Finding even one NS companion to a spectroscopically characterized LMWD would be very valuable, since this system would constrain the NS mass.
% TO ADD?: Also allows constraints on MSP age, braking index
However, radio and X-ray searches for NS companions to LMWDs have been unsuccessful \citep{agueros09b,agueros09a,kilic13}.

For each LMWD in the ELM sample, spectroscopy provides $\period$, the primary WD mass ($M_1$), and the projected orbital velocity ($K=v \sin i$). From Kepler's third law, we can write (assuming circular orbits):
\begin{equation}
	\frac{(M_2 \sin i)^3}{\left(M_1+M_2\right)^2} = \frac{\period}{2\pi G} K^3. \label{eq:massfunc}
\end{equation}
The righthand side is the mass function $\mf$. The companion mass, $M_2$, is minimized for an edge-on orbit ($i = 90\degree$). Because of this dependence on $i$, the nature of the companion cannot usually be determined based on $\mf$ alone. Figure~\ref{fig:Porb-M1} shows that the population of LMWDs with pulsar companions occupies the same region in $M_1 - \period$ space as those with WD companions. Therefore, barring rare circumstances (such as eclipsing systems), individual LMWDs with NS companions cannot be identified from optical observations alone.

The ELM sample is now large enough that $f_{\rm NS}$ and the companion mass distribution can be statistically constrained. These constraints provide important tests for population synthesis models.
%Observational constraints on $f_{\rm NS}$ and the WD companion mass distribution make LMWDs an important benchmark population for population synthesis models. The ELM sample is now large enough these parameters can be statistically constrained.
%Even if no NS companions to LMWDs are found, constraints on the WD companion mass distribution to LMWDs provide observations for population synthesis studies to match. The ELM sample is now large enough that the NS fraction and WD companion mass distribution can be statistically constrained. 
In this work, we develop a probabilistic model to infer parameters of an assumed form for the mass distribution of companions to LMWDs. Our method is similar to that employed by \citet{ozel12} and \citet{kiziltan13} to describe the mass distribution of NSs in binaries using post-Keplerian parameters. We focus on several specific questions: Can the population as a whole be modeled using a simple description of the companion masses? How does the companion mass distribution compare to predictions from population synthesis simulations?  What is the rate of NS-LMWD binaries implied by our model? What are the resulting distributions of NS probabilities for individual systems in the ELM sample?

To answer these questions, we build the mathematical framework (Section 2), then test our resulting model (Section 3). We apply our model to the ELM WD sample (Section 4) before concluding (Section 5).

\begin{figure}[h!]
\begin{center}
\includegraphics[angle=90,width=0.95\columnwidth]{f1.eps}
\caption{The $M_1$ - $\period$ distribution of the ELM WD sample (circles) and the known WD-NS binaries (triangles). The three eclipsing systems in the ELM sample with known $M_2$ are shown as filled circles, and the masses of the ELM WDs without detected radial velocity variations are shown by the arrows. From $M_1$ and $\period$ alone, the two populations are indistinguishable.}
\label{fig:Porb-M1}
\end{center}
\end{figure}

\section{Building our model}

We construct a statistical model to constrain a parametric model for the distribution of LMWD companion masses, $p(M_2 \given \bs{\theta})$.\footnote{We represent vectors or sets of parameters or quantities by bold symbols.} For each system, we assume we have $K$, $T$, and $M_1$, and therefore know the value of $\mf$ (\eqn~\ref{eq:massfunc}). Our model is based on a Bayesian approach, but this problem could also be viewed from a maximum-likelihood perspective. We wish to derive posterior constraints on the model parameters, $\bs{\theta}$, that describe the distribution of companion masses, $p(M_2\given \bs{\theta})$, given the set of observed mass functions, $\bs{m_f}$, by deconvolving the distribution of $\mf$'s from the unobserved $i$'s. Using Bayes' rule,
\begin{equation}
    p(\bs{\theta} \given \bs{\mf}) = \frac{1}{\mathcal{Z}}~p(\bs{\mf} \given \bs{\theta})~p(\bs{\theta}).
\end{equation}
where $p(\bs{\mf} \given \bs{\theta})$ is the likelihood, $p(\bs{\theta})$ is the prior on parameters $\bs{\theta}$, and the evidence integral, $\mathcal{Z}$, is a constant that depends only on the data. The likelihood, $p(\bs{\mf} \given \bs{\theta})$, can be split into a product over the likelihoods of individual systems:

\begin{equation}
p(\bs{\mf} \given \bs{\theta}) = \prod_j p(\mf \given \bs{\theta})
\end{equation}
where the product is over each of the $j$ systems.  This marginal likelihood involves integrals over the unobserved quantities $i$ and $M_2$,
\begin{align}
    p(\mf \given \bs{\theta}) &= \int_0^\infty dM_2 \int_0^{\pi/2} di  \nonumber \\
      & \qquad {} \times p(\mf \given M_1, M_2, i)~p(M_2 \given \bs{\theta})~p(i).
\end{align}
We neglect observational uncertainties in $\mf$ and $M_1$,\footnote{The fractional uncertainties in these quantities are small, $\sigma_x / x \sim 0.05-0.1$ \citep{gianninas14}.} and assume the inclination angles are isotropically distributed:
\begin{equation}
	p(\mf \given M_1, M_2, i) = \delta \left[\mf - f(M_1, M_2, i) \right]
\end{equation}
where
\begin{equation}
	f(M_1, M_2, i) = \frac{(M_2 \sin i)^3}{(M_1 + M_2)^2}
\end{equation}
and
\begin{equation}
	p(i) = \sin i.
\end{equation}
For now, we do not specify a parametric form for the companion mass distribution, $p(M_2 \given \bs{\theta})$. With the above assumptions, the marginal likelihood integral is:
\begin{align}
    p(\mf \given \bs{\theta}) &= \int_{0}^\infty dM_2 ~p(M_2 \given \bs{\theta})  \nonumber \\
    & \qquad {} \times \int_0^{\pi/2} di ~\sin i ~ \delta \left[g(M_1,M_2,i) \right]\label{eq:delta}
\end{align}
where
\begin{equation}
	g(M_1,M_2,i) = \mf - \frac{M_2^3}{(M_1+M_2)^2}\sin^3 i.
\end{equation}
The inner integral (over $i$) has the form:
\begin{equation}
    \int dx~F(x)~\delta \left[ G(x) \right] = \sum_j \frac{F(x^*_j)}{|G'(x^*_j)|}
\end{equation}
where the sum is over the roots, $x^*_j$, of the function $G(x)$. The root, $i^*$, and derivative of the argument of the delta function in \eqn\ref{eq:delta} are: 
\begin{align}
	\sin i^* &= \frac{ \left[\mf(M_1+M_2)^2 \right]^{1/3}}{M_2}\\
	\frac{\partial g}{\partial i}\bigg\rvert_{i^*} &= \frac{3M_2^3}{(M_1+M_2)^2} \sin^2 i^* \sqrt{1 - \sin^2 i^*}
\end{align}
We may rewrite the marginal likelihood as:
\begin{align}
	p(\mf \given \bs{\theta}) &= \int_{0}^\infty dM_2~p(M_2 \given \bs{\theta})~\sin i^* \left(\frac{\partial g}{\partial i}\bigg\rvert_{i^*}\right)^{-1}\\
	&= \int_{M_{2,{\rm min}}}^\infty dM_2~p(M_2 \given \bs{\theta})~h(M_2, \mf, M_1) \label{eq:fullm2}
\end{align}
where the bottom bound in the integral in \eqn\ref{eq:fullm2} is set by the minimum companion mass for which the integrand is real, $M_{2,{\rm min}}$ (determined by setting $i=90\degree$ in \eqn\ref{eq:massfunc} and solving for $M_2$), and

\begin{equation}
h(M_2, \mf, M_1) = \frac{(M_1+M_2)^{4/3}}{3\ \mf^{1/3}M_2\sqrt{M_2^2 - \left[ \mf(M_1+M_2)^2 \right]^{2/3}}}
\end{equation}

\subsection{Our Model} \label{sec:experiments}

>>>>

\title{The mass distribution of companions to low-mass white dwarfs}
\author{Jeff J.~Andrews\altaffilmark{\colum}, Adrian M.~Price-Whelan\altaffilmark{\colum}, Marcel A.~Ag\"ueros\altaffilmark{\colum}}

@@

\begin{abstract}
Measuring the masses of companions to single-line spectroscopic binary stars is (in general) not possible because of the unknown orbital plane inclination.  Even when the mass of the visible star can be measured, only a lower limit can be placed on the mass of the unseen companion. However, since these inclination angles should be isotropically distributed, for a large enough, unbiased sample, the  companion mass distribution can be deconvolved from the distribution of observables. In this work, we construct a hierarchical probabilistic model to infer properties of unseen companion stars given observations of the orbital period and projected radial velocity of the primary star. We apply this model to three mock samples of low-mass white dwarfs (LMWDs, $M\lesssim0.45~\Msun$) and a sample of post-common-envelope binaries. We use a mixture of two Gaussians to model the WD and neutron star (NS) companion mass distributions. Our model successfully recovers the initial parameters of these test data sets. We then apply our model to 55 WDs in the extremely low-mass (ELM) WD Survey. Our maximum a posteriori model for the WD companion population has a mean mass $\mu_{\rm WD} = 0.74~\Msun$, with a standard deviation $\sigma_{\rm WD} = 0.24~\Msun$. Our model constrains the NS companion fraction $f_{\rm NS}$ to be $<$16\% at 68\% confidence. We make samples from the posterior distribution publicly available so that future observational efforts may compute the NS probability for newly discovered LMWDs.
\end{abstract}

@@

\section{Introduction}
Except in cases of extreme metallicity \citep{kilic07}, the Galaxy is not old enough to produce low-mass white dwarfs (LMWDs) through single-star evolution. Instead, LMWDs are expected to form through interactions with another star \citep{han98,nelemans00,nelemans01,vdSluys06,woods12}. Indeed, with few exceptions, follow-up observations consistently find companions to LMWDs \citep{marsh95,maxted00,nelemans05,rebassa11}. Recently, the ELM WD Survey has identified 61 extremely LMWDs ($M\lesssim0.3~ \Msun$) in the Sloan Digital Sky Survey \citep[SDSS;][]{york00} and elsewhere \citep{ELMI,ELMII, ELMIII, ELMIV, ELMV}. We refer to the 55 WDs found by these authors that have a measured radial velocity (RV) and orbital period ($\period$) as the ELM sample.

These RV and $\period$ measurements indicate that the LMWDs companions are most likely WDs. However, since the inclination angle $i$ is unknown, LMWDs could have neutron star (NS) companions. Indeed, LMWDs are known companions to millisecond pulsars, although these WDs are generally too faint for spectroscopy \citep{vKerkwijk96,callanan98,bassa06,antoniadis12}. Finding even one NS companion to a spectroscopically characterized LMWD would be very valuable, since this system could constrain the NS mass. To date, unfortunately, radio and X-ray searches for NS companions to LMWDs have been unsuccessful \citep{vLeeuwen07,agueros09b,agueros09a,kilic13}.

For each LMWD in the ELM sample, spectroscopy provides $\period$, the primary WD mass $M_1$, and the projected orbital velocity $K=v \sin i$. Assuming circular orbits, we can write:
\begin{equation}
	\frac{(M_2 \sin i)^3}{\left(M_1+M_2\right)^2} = \frac{\period}{2\pi G} K^3, \label{eq:massfunc}
\end{equation}
where the right side is the mass function $\mf$. The companion mass, $M_2$, is minimized for an edge-on orbit ($i = 90\degree$). Because of this dependence on $i$, the nature of the companion cannot usually be determined based on $\mf$ alone. Figure~\ref{fig:Porb-M1} shows that the population of LMWDs with pulsar companions occupies the same region in $M_1 - \period$ space as those with WD companions. Therefore, barring rare circumstances such as eclipsing systems, individual LMWDs with NS companions cannot be identified from optical observations alone.

The ELM sample is now large enough that the $M_2$ distribution and NS companion fraction can be constrained statistically. %These constraints provide important tests for population synthesis models. 
We have developed a probabilistic model to infer parameters of an assumed form for the $M_2$ distribution. Our method is similar to that employed by \citet{ozel12} and \citet{kiziltan13} to describe the mass distribution of NSs in binaries using post-Keplerian parameters. We focus on the following questions: Can the companion population be modeled using a simple description of $M_2$? How does the $M_2$ distribution compare to predictions from population synthesis simulations?  What is the rate of LMWD-NS binaries implied by our model? What are the resulting distributions of NS probabilities for individual systems in the ELM sample?

To answer these questions, we build the mathematical framework (Section 2), then test our resulting model (Section 3). We apply our model to the ELM sample (Section 4) before concluding (Section 5).

\begin{figure}[h!]
\begin{center}
\includegraphics[angle=90,width=0.95\columnwidth]{f1.eps}
\caption{The $M_1$ - $\period$ distribution of the ELM sample (circles) and the known WD-NS binaries (triangles). The three eclipsing systems in the ELM sample with known $M_2$ are shown as filled circles, and the masses of the ELM WDs without detected RV variations are shown by the arrows. From $M_1$ and $\period$ alone, the two populations are indistinguishable.}
\label{fig:Porb-M1}
\end{center}
\end{figure}

\section{Building our model}
We construct a statistical model to constrain a parametric model for the distribution of LMWD companion masses, $p(M_2 \given \bs{\theta})$.\footnote{We represent vectors or sets of parameters or quantities by bold symbols.} For each system, we assume we have $K$, $T$, and $M_1$, and therefore know $\mf$. We wish to derive posterior constraints on the model parameters, $\bs{\theta}$, which describe the distribution of companion masses, $p(M_2\given \bs{\theta})$, given the set of observed mass functions, $\bs{m_f}$, by deconvolving the $\mf$ distribution from the unobserved inclinations. Using Bayes' rule,
\begin{equation}
    p(\bs{\theta} \given \bs{\mf}) = \frac{1}{\mathcal{Z}}~p(\bs{\mf} \given \bs{\theta})~p(\bs{\theta}),
\end{equation}
where $p(\bs{\mf} \given \bs{\theta})$ is the likelihood, $p(\bs{\theta})$ is the prior on parameters $\bs{\theta}$, and the evidence integral, $\mathcal{Z}$, is a constant that depends only on the data. The likelihood, $p(\bs{\mf} \given \bs{\theta})$, can be split into a product over the likelihoods of individual systems:
\begin{equation}
p(\bs{\mf} \given \bs{\theta}) = \prod_j p(\mf \given \bs{\theta}),
\end{equation}
where the product is over each of the $j$ systems.  This marginal likelihood involves integrals over the unobserved quantities $i$ and $M_2$,
\begin{align}
    p(\mf \given \bs{\theta}) &= \int_0^\infty dM_2 \int_0^{\pi/2} di  \nonumber \\
      & \qquad {} \times p(\mf \given M_1, M_2, i)~p(M_2 \given \bs{\theta})~p(i).
\end{align}
We neglect observational uncertainties in $\mf$ and $M_1$,\footnote{The fractional uncertainties in these quantities are small, $\sigma_x / x \sim 0.05-0.1$ \citep{gianninas14}.} and assume the inclination angles are isotropically distributed:
\begin{equation}
	p(\mf \given M_1, M_2, i) = \delta \left[\mf - f(M_1, M_2, i) \right],
\end{equation}
where
\begin{equation}
	f(M_1, M_2, i) = \frac{(M_2 \sin i)^3}{(M_1 + M_2)^2}
\end{equation}
and
\begin{equation}
p(i) = \sin i.
\end{equation}
For now, we do not specify a parametric form for the companion mass distribution, $p(M_2 \given \bs{\theta})$. With the above assumptions, the marginal likelihood integral is:
\begin{align}
    p(\mf \given \bs{\theta}) &= \int_{0}^\infty dM_2 ~p(M_2 \given \bs{\theta})  \nonumber \\
    & \qquad {} \times \int_0^{\pi/2} di ~\sin i ~ \delta \left[g(M_1,M_2,i) \right]\label{eq:delta},
\end{align}
where
\begin{equation}
	g(M_1,M_2,i) = \mf - \frac{M_2^3}{(M_1+M_2)^2}\sin^3 i.
\end{equation}
The inner integral (over $i$) has the form:
\begin{equation}
    \int dx~F(x)~\delta \left[ G(x) \right] = \sum_j \frac{F(x^*_j)}{|G'(x^*_j)|},
\end{equation}
where the sum is over the roots, $x^*_j$, of the function $G(x)$. The root, $i^*$, and derivative of the argument of the delta function in \eqn\ref{eq:delta} are: 
\begin{align}
	\sin i^* &= \frac{ \left[\mf(M_1+M_2)^2 \right]^{1/3}}{M_2}, \\
	\frac{\partial g}{\partial i}\bigg\rvert_{i^*} &= \frac{3M_2^3}{(M_1+M_2)^2} \sin^2 i^* \sqrt{1 - \sin^2 i^*}.
\end{align}
We may rewrite the marginal likelihood as:
\begin{align}
	p(\mf \given \bs{\theta}) &= \int_{0}^\infty dM_2~p(M_2 \given \bs{\theta})~\sin i^* \left(\frac{\partial g}{\partial i}\bigg\rvert_{i^*}\right)^{-1}\\
	&= \int_{M_{2,{\rm min}}}^\infty dM_2~p(M_2 \given \bs{\theta})~h(M_2, \mf, M_1). \label{eq:fullm2}
\end{align}
The bottom bound in the integral in \eqn\ref{eq:fullm2} is set by the minimum companion mass for which the integrand is real, $M_{2,{\rm min}}$, determined by setting $i=90\degree$ in \eqn\ref{eq:massfunc} and solving for $M_2$, and

\begin{equation}
h(M_2, \mf, M_1) = \frac{(M_1+M_2)^{4/3}}{3\ \mf^{1/3}M_2\sqrt{M_2^2 - \left[ \mf(M_1+M_2)^2 \right]^{2/3}}}.
\end{equation}

\subsection{Our Model} \label{sec:experiments}
We must now choose a functional form for the companion mass distribution, $p(M_2\given \bs{\theta})$. We use a two-component Gaussian mixture model. We truncate the distributions using physically motivated bounds: the WD component is restricted to $M_2\in [0.2,\wdupper]~\Msun$ and the NS component is restricted to $M_2\in [1.3,2.0]~\Msun$. We then have:
\begin{align}
	p(M_2 \given \bs{\theta}) &= \left[ (1-f_{\rm NS})~p_{\rm WD} + f_{\rm NS}~p_{\rm NS} \right], 
\end{align}
where $f_{\rm NS}$ is the NS fraction and
\begin{align}
	p_{\rm WD} &= \mathcal{N}(M_2 \given \mu_{\rm WD}, \sigma^2_{\rm WD}); ~0.2 < \frac{M_2}{\Msun} < \wdupper, \\
	p_{\rm NS} &= \mathcal{N}(M_2 \given \mu_{\rm NS}, \sigma^2_{\rm NS}); ~1.3 < \frac{M_2}{\Msun} < 2.
\end{align}
$\mathcal{N}$ is the (truncated, but properly normalized) normal distribution with mean $\mu$, variance $\sigma^2$; the distributions are limited to the ranges specified. To reduce the number of parameters in our model we fix $\mu_{\rm NS}$ and $\sigma_{\rm NS}$
%$\mu_{\rm NS} = 1.4~\Msun$ and $\sigma_{\rm NS} = 0.05~\Msun$, 
to:
\begin{align}
	\mu_{\rm NS} &= 1.4~\Msun, \\
	\sigma_{\rm NS} &= 0.05~\Msun,
\end{align}
as some NSs in binaries may be somewhat more massive than the canonical NS mass of 1.35 \Msun~\citep{kiziltan13,smedley14}.

The probability of any particular WD having a NS companion, $P_{\rm NS}$, can be computed for a given set of parameters for the $M_2$ distribution:
\begin{equation}
P_{\rm NS} = \frac{\int_{M_{2,{\rm min}}}^{\infty} dM_2~ f_{\rm NS}~ p_{\rm NS}~ h(M_2, \mf, M_1)}{p(\mf \given \bs{\theta})}. \label{eq:P_NS}
\end{equation}

Our companion mass model parameters are then $\bs{\theta} = (\mu_{\rm WD}, \sigma_{\rm WD}, f_{\rm NS})$. For $\mu_{\rm WD}$, we use a uniform prior from $0.2-1.0~\Msun$; for $\sigma_{\rm WD}$, we use a logarithmic (scale-invariant) prior over the range $0.02-2.0~\Msun$. Finally, we use a uniform prior over the dimensionless $f_{\rm NS}$ from $0-1$. The model parameters are summarized in Table~\ref{tbl:parameters}.
